\begin{thebibliography}{52}
\bibitem{AKP}
Alexander, S., Kapovitch, V., Petrunin, A.,
\textit{Alexandrov geometry---foundations.}
Grad. Stud. Math., 236 (2024).

\bibitem{BBI} Burago, D.; Burago, Yu.; Ivanov, S., \textit{A course in metric geometry.} Grad. Stud. Math., 33. (2001).

\bibitem{gromov-CAT} Gromov, M., \textit{CAT($\kappa$)-spaces: Construction and Concentration,} Journal of Mathematical Sciences, Vol. 119, No. 2, (2004) 178--200.

\bibitem{isbell} Isbell, J. \textit{Six theorems about metric spaces}. Comment. Math. Helv. 39: 65--74.(1964).

\bibitem{kirszbraun} Kirszbraun, M. D., \textit{\"Uber die zusammenziehende und Lipschitzsche Transformationen,} Fundamenta Math. 22 (1934), 77--108.

\bibitem{kleiner} Kleiner, B.,  \textit{The local structure of length spaces
with curvature bounded above,} Math. Z. 231, 409--456 (1999).

\bibitem{kleiner-leeb}  Kleiner, B.; Leeb, B., \textit{Rigidity of quasi-isometries for symmetric spaces and Euclidean buildings.} Inst. Hautes \'Etudes Sci. Publ. Math. No. 86 (1997), 115--197 (1998).

\bibitem{lang-schroeder}  Lang, U.; Schroeder, V., \textit{Kirszbraun's theorem and metric spaces of bounded curvature,} Geom. funct. anal.
Vol. 7 (1997) 535--560.

\bibitem{LPZ} Lebedeva, N.;
Petrunin, A.; Zolotov, V., \textit{Bipolar comparison.}
Geom. Funct. Anal. 29 (2019), no. 1, 258--282.


\bibitem{monod} Monod, N. \textit{Superrigidity for irreducible lattices and geometric splitting,} J. of the AMS,
Vol. 19, Num. 4, October 2006, 781--814.

\bibitem{plaut:survey}
Plaut, C., \textit{Metric spaces of curvature $\ge k$.}
Handbook of geometric topology,  819--898, North-Holland, Amsterdam, 2002.

\bibitem{reshetnyak:major}
Reshetnyak, Yu. G.,
\textit{Nonexpanding maps in a space of curvature no greater than $K$},
Sibirskii Mat. Zh. 9 (1968), 918-928 (Russian).
English translation: \textit{Inextensible mappings in a space of curvature
no greater than $K$}, Siberian Math. Jour. 9 (1968),683-689.


\bibitem{sturm}
Sturm, K. T. \textit{Metric spaces of lower bounded curvature.}
Exposition. Math.  17  (1999),  no. 1, 35--47.

\bibitem{valentine-sphere}
Valentine F.A., Contractions in non-Euclidean spaces, Bull. Amer. Math. Soc. 50 (1944), 710--713.

\bibitem{valentine-kirszbraun}
Valentine, F. A., A Lipschitz condition preserving extension for a vector
function, Amer. J. Math. 67 (1945), 83--93.

\bibitem{vilms}
Vilms, J., \textit{Totally geodesic maps}.
J. Differential Geometry 4 1970 73--79.
\end{thebibliography}
